---
title: "EXCAM Campylobacter qPCR analysis"
author: "Arie Havelaar, Yang Yang"
format: html
echo: false
bibliography: references.bib
---

## Introduction

This script provides a descriptive analysis of qPCR data from the EXCAM project and fits censored distributions to the data for use in exposure analysis using Bayesian estimation in `Stan`. qPCR data are available with PMAxx treatment, representing DNA in viable cells and total DNA of *Campylobacter*, respectively. After many discussions with the lab team, we decided to only fit the viability qPCR data and not to compare them with direct PCR because there were too many differences in handling of the samples. Also, no samples were tested for total DNA from sources ED and ES. Even within the dataset for viability qPCR, there are differences in handling. One important reason is the difference in total DNA concentration between sample types, which necessitated pre-concentration of some samples. We have no information on the impact of these differences in sample handling on DNA concentrations in the final test samples so cannot account for these differences. As recoveries are by definition less than 100%, it is safe to assume that the DNA concentrations are underestimated more if more sample preparation steps are included.

## Housekeeping

```{r}
#| output: false
library(fda)
library(tidyverse)
library(ggpubr)
library(gt)
library(tidyverse)
library(rstan)
library(bayesplot)
library(patchwork)
library(survival)
library(modelr)
library(lpridge)
library(extraDistr)

#library(conflicted)
#conflict_prefer("select", "dplyr")
#conflicts_prefer(dplyr::filter)
#conflicts_prefer(MASS::area)
#conflicts_prefer(tidyr::complete)
#conflicts_prefer(rstan::extract)
#conflicts_prefer(fda::gait)

# Stan options
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)

```

## Data

Data are available as rds files after processing the original lab created Excel files using the script `20250812_data_processing.qmd`.

```{r}
df_via <- readRDS("data/df_via.rds") # Viable DNA
```

### Count proportion of censored observations

Due to technical limitations in the qPCR test, a Limit of Detection (LOD) is defined as Ct \> 35. For soil samples, the LOD is defined as Ct \> 29. Samples with higher Ct values are considered censored. This does not necessarily mean they are "negative". Instead, the concentration of DNA was too low to measure reliably. For this reason, we avoid the concept of prevalence in the data analysis and prefer to fit censored distributions.

```{r}
# Number of samples by source, viable DNA
df_via_n <- 
  df_via |> 
    group_by(source)  |> 
      summarize(n())
colnames(df_via_n) <- c("source", "n_via")

# Number of censored samples by source, viable DNA
df_via_cens <- 
  df_via |> 
    filter(censored == 1) |> 
    group_by(source)  |> 
      summarize(n())
colnames(df_via_cens) <- c("source", "cens_via")

# Join all data in one dataframe

df_cens_count <- 
df_via_n |> 
  left_join(df_via_cens, by = "source")

df_cens_count$prop_via <- with(df_cens_count, cens_via / n_via)

lvl<- levels(df_via$source) # Levels for viability data

df_samples <- data.frame(
  source_code = lvl,
  source_name = c("Bathing water", "Drinking water", "Food", "Soil",
                  "Fomite", "Areola swab", "Breast milk",
                  "Sibling hand rinse", "Mother hand rinse", "Infant hand rinse") 
)

df_samples$via <- df_via_n$n_via
df_samples$cens_via <- df_via_cens$cens_via
df_samples$prop_via <- round(with(df_cens_count, cens_via / n_via), 2)

df_samples |> 
  gt() |> 
  cols_label(
    source_code = "Source code",
    source_name = "Source",
    via = "Total samples",
    cens_via = "Censored samples",
    prop_via = "Proportion censored"
  )
```

### Plot censoring by source and method

```{r}
#| warning: false

# Prepare dataframe
df_cens_plot <-  df_samples[, c(1, 4)]
df_cens_plot <-df_cens_plot |> 
      reshape2::melt()
colnames(df_cens_plot) <- c("source", "DNA", "value")

(df_cens_plot |> 
  ggplot(aes(x = value, y = source)) +
    geom_point(color = "#0072B2") +
    scale_x_continuous(labels = scales::percent) +
    scale_y_discrete(labels = c("EB" = "Bathing water",
                                "ED" = "Drinking water",
                                "EF" = "Food",
                                "ES" = "Soil",
                                "EV" = "Fomites",
                                "HA" = "Areola swab",
                                "HB" = "Breast milk",
                                "HH" = "Sibling hand rinse",
                                "HP" = "Mother hand rinse",
                                "HR" =  "Infant hand rinse"),
                    limits = rev) +
    labs(x = "Proportion censored observations",
         y = "Source") +
    theme_light()
)
#ggsave("../manuscript/cens_plot.tiff")
```

## Fit censored normal distributions

We fit censored distributions using the `rstan` package. The censored model is defined in the file `norm_cens_model.stan`.

```{r}
# Define the model
norm_cens_model <- stan_model(file = "Stan/norm_cens_model.stan")
t_cens_model <- stan_model(file = "Stan/t_cens_model.stan")
laplace_cens_model <- stan_model(file = "Stan/Laplace_cens_model.stan")
```

### Source EB - bathing water

`r round(df_samples[which(df_samples$source_code == "EB"), 5] * 100, 0)`% censored data.

```{r}
#| output: true
#| warning: true

extract_data<-function(source_name){
   y <- df_via |> 
     filter(source == source_name & !is.na(dna)) |> 
     select(c(dna)) |> 
     unlist()
   lod <- df_via |> 
     filter(source == source_name & !is.na(dna)) |> 
     select(c(dna_lod)) |> 
     unlist()
   cen <- df_via |> 
     filter(source == source_name & !is.na(dna)) |> 
     select(c(censored)) |> 
     unlist()
   list(y=y, lod=lod, cen=cen, N = length(y))
}

data_tot_EB <- extract_data('EB')

# Fit the model
fit_tot_EB <- sampling(norm_cens_model, data = data_tot_EB, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_EB, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()
  

# check trace plot
traceplot(fit_tot_EB, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_EB)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2


# Empirical CDF
get_empirical_cdf <- function(dat) {
   data_cens <- data.frame(left=rep(NA, length(dat$y)), right=rep(NA, length(dat$y))) 
   data_cens$left <- ifelse(dat$cen==1, NA, dat$y)
   data_cens$right <- ifelse(dat$cen==1, dat$lod, dat$y)

   Surv_obj <- Surv(time = data_cens$left,
                 time2 = data_cens$right,
                 type = "interval2")

   fit_surv <- survfit(Surv_obj ~ 0)

    data.frame(
     x = fit_surv$time,
     cdf  = 1 - fit_surv$surv,
     sd = fit_surv$std.err)
}

empirical_cdf <- get_empirical_cdf(data_tot_EB)

#locpol <-lpridge(empirical_cdf$x, empirical_cdf$cdf, bandwidth=floor(nrow(empirical_cdf)*0.50), deriv=0, n.out=100, x.out=seq(0, 7, length.out=100), order = 3, ridge = 0, weight = "epa")
#smooth.emp_cdf<-data.frame(x=locpol$x.out, cdf=locpol$est) 
#smooth.emp_cdf$cdf[which(smooth.emp_cdf$cdf<0)] <- 0
#smooth.emp_cdf$cdf[which(smooth.emp_cdf$cdf>1)] <- 1
#k<-nrow(smooth.emp_cdf)
#y <- c(smooth.emp_cdf$cdf[1], smooth.emp_cdf$cdf[2:k] - smooth.emp_cdf$cdf[1:(k-1)])
#y[y<0] <- 0
#smooth.emp_cdf$cdf <- cumsum(y)

# monotone-function smoothing
monosmoothing <- function(emp)
{
   #wgt=1/(emp$sd**2)
   norder <- 6
   Lfdobj=2
   lambda=10^(0.5)
   rng <- range(emp$x)
   ntime   <- length(emp$x)
   nbasis <- ntime + norder - 2
   wbasis <- create.bspline.basis(rng, nbasis, norder, emp$x)
   cvec0 <- matrix(0,nbasis,1) #  starting values for coefficient
   Wfd0  <- fd(cvec0, wbasis)
   result <- smooth.monotone(empirical_cdf$x, empirical_cdf$cdf, fdPar(Wfd0, Lfdobj, lambda),     dbglev=0)
   Wfd  <- result$Wfdobj
   beta <- result$beta
   #  Evaluate the fitted height curve over a fine mesh
   x.new <- seq(0, 7, length.out=100)
   y.new <- beta[1] + beta[2]*eval.monfd(x.new, Wfd)
   y.new[y.new<0] <- 0
   y.new[y.new>1] <- 1
   data.frame(x=x.new, cdf=y.new) 
}
smooth.emp_cdf<-monosmoothing(empirical_cdf)
  
#plot(monosmooth.emp_cdf$x, monosmooth.emp_cdf$cdf, type='p', pch = 19, cex = 0.6)
#lines(empirical_cdf$x, empirical_cdf$cdf, lty=2, lwd=2, col='red')

empirical_quantile <- function(p, emp) {
  sapply(p, function(pp) {
                  idx <- which(emp$cdf >= pp)[1]
                  emp$x[idx]
            })
}
#p <- seq(0.01, 0.99, by = 0.01)
p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles
# Theoretical CDF
post_sum <- summary(fit_tot_EB)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
#nu_hat <- post_sum["nu", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)
#theoretical_q <- mu_hat + sigma_hat * qt(p, nu_hat) 
#theoretical_q <- qlaplace(p, mu_hat, sigma_hat)
#sum(abs(theoretical_q-empirical_q1))
#sum(abs(theoretical_q-empirical_q2))
#sum(abs(theoretical_q-empirical_q3))

integral_trapz <- function(x, y) {
  idx <- 2:length(x)
  # The formula is the sum of areas of trapezoids: 0.5 * base_width * (height1 + height2)
  return(as.double((x[idx] - x[idx-1]) %*% (y[idx] + y[idx-1])) / 2)
}

# check the mean calculated from smoothed empirical CDF
integral_trapz(smooth.emp_cdf$x, 1-smooth.emp_cdf$cdf)


# Plot empirical and theoretical CDF
plot(theoretical_q, empirical_q2, xlim=c(0, 7),ylim=c(0, 7),
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

### Source ED - drinking water

`r round(df_samples[which(df_samples$source_code == "ED"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true

# Provide the data as a list
data_tot_ED <- extract_data('ED')

# Fit the model
fit_tot_ED <- sampling(norm_cens_model, data = data_tot_ED, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_ED, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_ED, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_ED)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2


empirical_cdf <- get_empirical_cdf(data_tot_ED)
monosmooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf)
empirical_q2 <- empirical_quantile(p, monosmooth.emp_cdf)

# Theoretical CDF
post_sum <- summary(fit_tot_ED)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

# Plot empirical and theoretical CDF
plot(theoretical_q, empirical_q2, xlim=c(0, 7),ylim=c(0, 7),
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))
```

### Source EF - food

`r round(df_samples[which(df_samples$source_code == "EF"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true


# Provide the data as a list
data_tot_EF <- extract_data('EF')

# Fit the model
fit_tot_EF <- sampling(norm_cens_model, data = data_tot_EF, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_EF, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_EF, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_EF)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2


empirical_cdf <- get_empirical_cdf(data_tot_EF)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_EF)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

### Source ES - soil

`r round(df_samples[which(df_samples$source_code == "ES"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true

# Provide the data as a list
data_tot_ES <- extract_data('ES')

# Fit the model
fit_tot_ES <- sampling(norm_cens_model, data = data_tot_ES, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_ES, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_ES, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_ES)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2

empirical_cdf <- get_empirical_cdf(data_tot_ES)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_ES)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

### Source EV - fomites

`r round(df_samples[which(df_samples$source_code == "EV"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true

# Provide the data as a list
data_tot_EV <- extract_data('EV')

# Fit the model
fit_tot_EV <- sampling(norm_cens_model, data = data_tot_EV, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_EV, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_EV, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_EV)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2

empirical_cdf <- get_empirical_cdf(data_tot_EV)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_EV)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

### Source HA - areola swab

`r round(df_samples[which(df_samples$source_code == "HA"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true

# Provide the data as a list
data_tot_HA <- extract_data('HA')

# Fit the model
fit_tot_HA <- sampling(norm_cens_model, data = data_tot_HA, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_HA, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_HA, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_HA)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2

empirical_cdf <- get_empirical_cdf(data_tot_HA)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_HA)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

### Source HB - breast milk

`r round(df_samples[which(df_samples$source_code == "HB"), 5] * 100, 0)`% censored data, so no data for fitting

### Source HH - Sibling hand rinse

`r round(df_samples[which(df_samples$source_code == "HH"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true


# Provide the data as a list
data_tot_HH <- extract_data('HH')

# Fit the model
fit_tot_HH <- sampling(norm_cens_model, data = data_tot_HH, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_HH, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_HH, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_HH)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2

empirical_cdf <- get_empirical_cdf(data_tot_HH)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_HH)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```
### Source HP - mother hand rinse

`r round(df_samples[which(df_samples$source_code == "HP"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true

# Provide the data as a list
data_tot_HP <- extract_data('HP')

# Fit the model
fit_tot_HP <- sampling(norm_cens_model, data = data_tot_HP, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_HP, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_HP, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_HP)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2

empirical_cdf <- get_empirical_cdf(data_tot_HP)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_HP)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

### Source HR - infant hand rinse

`r round(df_samples[which(df_samples$source_code == "HR"), 5] * 100, 0)`% censored data.

```{r}

#| warning: true

# Provide the data as a list
data_tot_HR <- extract_data('HR')

# Fit the model
fit_tot_HR <- sampling(norm_cens_model, data = data_tot_HR, chains = 4, 
                iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
summary(fit_tot_HR, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
  as.data.frame() |> 
  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
  gt()

# check trace plot
traceplot(fit_tot_HR, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
posterior <- as.array(fit_tot_HR)
color_scheme_set("brightblue")
p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
p1 + p2

empirical_cdf <- get_empirical_cdf(data_tot_HR)
smooth.emp_cdf<-monosmoothing(empirical_cdf)

p <- empirical_cdf$cdf[empirical_cdf$cdf > 0 & empirical_cdf$cdf < 1]
empirical_q1 <- empirical_quantile(p, empirical_cdf) # non-smoothed empirical quantiles
empirical_q2 <- empirical_quantile(p, smooth.emp_cdf) # smoothed empirical quantiles

# Theoretical CDF
post_sum <- summary(fit_tot_HR)$summary
mu_hat <- post_sum["mu", "50%"]
sigma_hat <- post_sum["sigma", "50%"]
#nu_hat <- post_sum["nu", "50%"]
#theoretical_q <- mu_hat + sigma_hat * qt(p, nu_hat) 
theoretical_q <- qnorm(p, mu_hat, sigma_hat)

x.range <- y.range <- range(c(theoretical_q, empirical_q1, empirical_q2), na.rm=TRUE)
plot(theoretical_q, empirical_q2, xlim=x.range, ylim=y.range,
     xlab = "Theoretical quantiles",
     ylab = "Sample quantiles",
     main = "Q-Q plot",
     pch = 19, cex = 0.8)
points(theoretical_q, empirical_q1, pch=21, cex=0.8, col='green')
abline(0, 1, col = "red", lty=2, lwd = 1)
legend('topleft', c('Non-smoothed', 'Smoothed'), pch=c(19,21), col=c('black','green'))

```

## Fit all distributions in one model with sample type as covariate

Based on results of single fits, data for food, fomites and areola swabs will be excluded because of a high percentage censoring.

```{r}
df_via_f <- df_via |> 
  filter(!source  %in% c("EF", "EV", "HB")) |> 
  droplevels()
```

Stan requires explicit dummy variables, the design matrix can be created with `modelr::model_matrix()`.

```{r}
X <- select(df_via_f,
       source, dna, dna_lod,censored) |> 
  na.omit() |> # Drop incomplete cases
  model_matrix(~ source) |>
    as.matrix()
```

Fit multivariate model to filtered data

```{r}

#| warning: true

#y <- df_via_f |> 
#  select(c(dna)) |> 
#  unlist()
#lod <- df_via |> 
#  select(c(dna_lod)) |> 
#  unlist()
#cen <- df_via |> 
#  select(c(censored)) |> 
#  unlist()


# Provide the data as a list
#data_via_all <- list(y=y, lod=lod, cen=cen, N = length(y))

# Fit the model
#fit_via_all <- sampling(norm_cens_model, data = data_via_all, chains = 4, 
 #               iter = 2000, warmup = 1000)

# Summarize posterior, including R_hat (Gelman-Rubin diagnostic)
#summary(fit_via_all, pars = c("mu", "sigma"), probs = c(0.025, 0.5, 0.975))$summary |> 
#  as.data.frame() |> 
#  mutate(across(where(is.numeric), \(x) round(x, digits = 2))) |> 
#  gt()

# check trace plot
#traceplot(fit_via_all, pars = c("mu", "sigma"), inc_warmup = FALSE)

# Posterior Density Plots
#posterior <- as.array(fit_via_all)
#color_scheme_set("brightblue")
#p1 <- mcmc_areas(posterior, pars = "mu", prob=0.9, area_method ="equal height")
#p2 <- mcmc_areas(posterior, pars = "sigma", prob=0.9, area_method ="equal height")
#p1 + p2
```